Automatic Matching of Bullet Lands

### 2016-04-18 revision/response.Rmd

***Yes, it would be possible to stress the classification system by using only parts of the profile, but a systematic evaluation of this by far exceeds the scope of this paper.***

11. What is the standard error of the error rates presented for the matching algorithms? The number of cross comparisons gives false impression the standard errors the when presenting the results.

***HH: I'm sure I am misunderstanding the comment, but I don't know how to answer this question. The standard error of a rate in (0,1) would be sqrt(p*(1-p)). So because p is small the square root is going to be still pretty small. We are presenting out of bag error rates for the random forest - maybe we need to explain what out-of bag means in more detail?***

***EH: Yeah, I was a little unsure of this. Maybe we can ask Alicia, and if she also is confused by it I can type up an explanation on how the OOB error is computed.***

@@

21. First, the term "land" is used in the title but not defined until p. 3.  I recommend changing the title, to ensure broad comprehenson; perhaps "Automatic Matching of Bullet Marks," or something of that kind.

***HH: Let's discuss that on Monday***

22. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth.

***EH: Presumably the typical width is some kind of class characteristic dependent on the gun itself, right?***

@@

23. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.

***DONE***

24. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.

***EH: We could slightly rotate the image in RGL, but the top-down view is nice. Alternatively I could just disable the axis labels - We don't really need to see them in this image, right?***

@@

25. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.

***EH: Hm. I think we phrased this correctly and the reviewer misunderstood - We literally ARE showing groove to groove impressions. Maybe we need to be more clear about groove impressions... but in the following paragraph + the image of the bullet I think its fairly clear. We could add in a little disclaimer like "Grooves, not to be confused with striae, are..."***

@@

48. Line 246: Replace ``this representations" with ``this representation" or ``these representations"

***DONE***

49. Caption of Figure 8: ``At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures." is not a complete sentence.

***HH: would a removal of the comma make it a complete sentence?***

***EH: I actually thought it already was a complete sentence... but I changed it to "The correlation peaks at a lag of -17, which indicates the largest amount of agreement between the signatures."***

### 2016-04-19 revision/response.Rmd

***Additional papers considered (not complete, and not yet included in the paper, but I think all of them should be cited): 
chumbley:2010 
uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening)
chubley:2010 is a much more detailed version over David Faden's thesis***

***chu:2010
describe extracting mean profile of an area of the image after untwisting the image ***

***bachrach:2002
general description of an automated matching system based on 3D imaging. 
Uses cross-correlation function for alignment.***

***bachrach:2006
67page final report for NIJ grant (for Intelligent Automation Inc.)
wear study: "difference between comparing the nth and (n+1)th bullet fired by a barrel as opposed to the nth and (n+100)th bullet fired by a barrel? In other words, does the proximity between firings make a difference? Although some wear effect could be detected in both of these scenarios, these effects do not appear to be significant enough to prevent correct identification"***

@@

***petraco:2012
I am not sure that this book is the same as the report for the NIJ grant. This is a huge effort, and we should certainly cite it, but there are not a lot of actual results included or comparisons done.***

***riva:2014
automatic matching system based on 3d images of cartridge cases. Matching involves breech faces, but no striation marks.***

***xie:2009
first sentence: "Every firearm has individual characteristics that are as unique to it as fingerprints are to human beings." the paper itself doesn't back up any of these claims. very disappointing***

***leon:2006
first sentence: "Fired bullets bear striation marks that can be thought of as a ‘‘fingerprint’’ left by the firearm." based on multiple 2d images. They extract a 1d 'signature'***

@@

Reviewer C:

14. there is a lack of knowledge of the existing literature on statistical work performed on tool marks and firearms. For example, the work performed by Nicholas Petraco, Fabiano Riva, or Alan Zheng in the past 5 years is not mentioned. Instead, old research projects are mentioned, which mostly focus on manual identification. The authors need to discuss the novelty of their method (and how it potentially addresses the shortcomings of these recent methods)

15.  The authors would need to study (or at least discuss extensively) the ability of their method to capture the variability in bullet profiles over time (at interval of several dozen of bullets fired by the same weapon, when the weapon has been cleaned, or when the bullet was degraded by hitting a target). In addition, a dataset of 10 firearms (even consecutively manufactured) does not provide any information on the rarity of a particular bullet profile, and therefore does not address the second probability statement. Alternatively, the proposed metrics could be tested in a database searching situation, which has not been done.

***At the time of the submission of this manuscript, the 35 bullets from Hamby set 252 were the only ones accessible to the public. They are part of Dr Zheng's efforts at NIST to provide a ballistics reference database. In the future more results will be available, but until then we are unable to provide a data-informed discussion of the very important issues raised by the reviewer.***

@@

21. First, the term "land" is used in the title but not defined until p. 3.  I recommend changing the title, to ensure broad comprehenson; perhaps "Automatic Matching of Bullet Marks," or something of that kind.

***HH: Let's discuss that on Monday***

22. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth.

***EH: Presumably the typical width is some kind of class characteristic dependent on the gun itself, right?***
***HH: I am not sure about the width of a striation - the width depends on where it is measured (in both x and z direction). I'm sure there are definitions of that available. We might need some more literature.***

@@

23. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.

***DONE***

24. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.

***EH: We could slightly rotate the image in RGL, but the top-down view is nice. Alternatively I could just disable the axis labels - We don't really need to see them in this image, right?***
***HH: I just deleted the values from the axis***

### 2016-04-19 revision/response.Rmd

***Additional papers considered (not complete, and not yet included in the paper, but I think all of them should be cited): 
chumbley:2010 
uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening)
chubley:2010 is a much more detailed version over David Faden's thesis***

***chu:2010
describe extracting mean profile of an area of the image after untwisting the image ***

***bachrach:2002
general description of an automated matching system based on 3D imaging. 
Uses cross-correlation function for alignment.***

***vorburger:2011
different uses of the cross-correlation function. One of them aligning profiles***

@@

24. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.

***EH: We could slightly rotate the image in RGL, but the top-down view is nice. Alternatively I could just disable the axis labels - We don't really need to see them in this image, right?***
***HH: I just deleted the values from the axis***

25. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.

***HH: this is clearly a misunderstanding, so we need to be nice about breaking this to the reviewer, and emphasize it in the paper some more.***
***EH: Hm. I think we phrased this correctly and the reviewer misunderstood - We literally ARE showing groove to groove impressions. Maybe we need to be more clear about groove impressions... but in the following paragraph + the image of the bullet I think its fairly clear. We could add in a little disclaimer like "Grooves, not to be confused with striae, are..."***

@@

49. Caption of Figure 8: ``At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures." is not a complete sentence.

***DONE***

***HH: would a removal of the comma make it a complete sentence?***

***EH: I actually thought it already was a complete sentence... but I changed it to "The correlation peaks at a lag of -17, which indicates the largest amount of agreement between the signatures."***
***HH: Ta, as they say around here :)***

### 2016-04-19 imaging-paper.Rnw
vorburger:2011 reference for ccf for alignment

\subsection{Horizontal alignment}
Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$. Striation marks show up in these representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lof$bulletname <- scrubPath(lof$bullet)
qplot(y, resid, group=bulletname, colour=bulletname, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lofY$bulletname <- scrubPath(lofY$bullet)
qplot(y+min(lof$y), resid, group=bulletname, colour=bulletname, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 taken at  heights of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

>>>>

\subsection{Horizontal alignment}
Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$. Striation marks show up in these representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lof$bulletname <- scrubPath(lof$bullet)
qplot(y, resid, group=bulletname, colour=bulletname, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lofY$bulletname <- scrubPath(lofY$bullet)
qplot(y+min(lof$y), resid, group=bulletname, colour=bulletname, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 taken at  heights of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
\hh{For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures~\cite{bachrach:2002, vorburger:2011, thompson:2013}.}
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).
